\section{Simulation code}

The class AliSimulation manages this part. An example is here : ``\$ALICE\_ROOT/EMCAL/
macros/TestEMCALSimulation.C''. The simulation
consists of different steps: geometry and event definition, particle
generation, transport of the particle in the material (GEANT) and
finally digitization. Note that the final output from the digitization
process is different from the processing of real experimental Raw Data. The process
of converting the digitized data to Raw Data is discussed in Sec.~\ref{sec:digi}.
Sec.~\ref{sec:simu_steps} gives the recipe to do all the steps of the simulation.


%\subsection{Event generation and particle transport: Hits}


%Once the generator is executed, the generated particles are transported
%in the detector material with the Monte Carlo code, GEANT3 by default. Other options are 
%GEANT4 or FLUKA\footnote{There may be some license problems  with FLUKA right now which could explain why it cannot be used at the moment}. All the generated particles are kept in a file called \textbf{Kinematics.root}. After the particle transport is executed, the objects \textbf{Hits}
%are created. They contain the energy deposited in the sensitive material
%of the detector by the generated particle, their position, impact
%time (after collision) and the identity of the original particle.
%Hits are stored in a file called \textbf{DETECTOR.Hits.root}, in the
%calorimeter case: \textbf{EMCAL.Hits.root.}

\input{StepManagerParticleTransport}
\clearpage


\subsection{Digitization: SDigits and Digits - Evi \label{sec:digi}}

We want to generate events which look like the real data collected
by the experiment. In the end, we want to have an amplitude in ADC
counts and a time (when particle traverse a cell) per each cell (tower)
of the calorimeter. In the code for calorimeters, it is done in the
following steps: 
\begin{enumerate}
\item \textbf{SDigit} objects are created, they consist
of the sum of deposited energy by all Hits in a cell (a particle can
create Hits in different cells but only one in a single cell), so
there is only one SDigit per fired cell.
\item \textbf{Digit} objects are created, they are like the SDigits but the energy in the cell
is transformed into the ADC amplitude units, the electronic noise
is added and Digits whose energy does not pass an energy threshold
(3 ADC counts) are eliminated. SDigits and Digits are stored in the files
\textbf{EMCAL.SDigits.root} and \textbf{EMCAL.Digits.root}, respectively.
\end{enumerate}

\subsection{Raw data - David \label{sec:simu_raw}}

The experiment does not record Digits directly, but instead a series of so-called
time samples with 10-bit ADC counts per channel. Each time bin is 100 ns
wide, corresponding to a 10 MHz readout.
These samples are referred to as
\textbf{Raw Data}. The samples follow a certain signal shape, more complicated than
a Gaussian distribution, which is fitted offline. 
The simulated signal (Gamma-2) shape is described in the AliEMCALRawResponse class,
in the RawResponseFunction method.
With real data, which is zero-suppressed, i.e. has the pedestal subtracted online, the
Digits amplitude is just the maximum of the distribution obtained
with the fit to the sample. The Digit time (defined by the time the
particle hits the active volume of the detector) is the time value at
the maximum signal fit. There are methods to go from Digits to
Raw and vice versa in the AliEMCALRawUtils class: Raw2Digits and Digits2Raw,
respectively. For the reconstruction step Digits are needed. The
generation of Raw Data is optional during simulations and the generated 
data can be reconstructed directly from Digits, but Raw data is the initial
step when reconstructing real data.


\subsection{How to make a simulation\label{sec:simu_steps}}

TestEMCALSimulation.C is a very simple macro where we specify all the simulation parameters
and proccess the simulation. Below is a similar but a bit more elaborated macro:



\begin{DDbox}{\linewidth}
\begin{lstlisting}
void TestEMCALSimulation() {

TString detector=``EMCAL TPC''; // Define in this variable the detectors that you want to be included in the simulation for the digitization. They can be less detectors than the detectors defined in the Config.C file, imagine that you want all the detectors in front of EMCal present to consider the conversion of particles but you are not really interested in the output from these detectors. 
// Option detector=``ALL''  makes all detectors. 

AliSimulation sim ; //Create simulation object 

// Generation and simulation 

sim.SetRunGeneration(kTRUE) ; //Default value is kTRUE, make generation  
// For some reason we may want to redo the Digitization, without redoing the generation, in this case it must set to kFALSE 

// Making SDigits 
sim.SetMakeSDigits(detector) ; //We want to make SDigits
// set no detectors if SDigits are already made

// Making Digits 
sim.SetMakeDigits(detector) ; //We want to make Digits 
// set no detectors if SDigits are already made

//Merging 
//sim.MergeWith(``bgrd/galice.root'') ; //If we want to merge a signal and a background, the merging is done at the SDigit level. The background must be located in the repertory defined in the method. 

//Write Raw Data, make Raw data from digits 
//sim.SetWriteRawData(detector) ;
//sim.SetConfigFile(``somewhere/ConfigXXX.C'');//Default is Config.C

//Make the simulation 
sim.Run(3) ; // Run the simulation and make 3 events 



\end{lstlisting}
\end{DDbox}