%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                             %
%        DISTRIBUTED ANALYSIS NOTE                            %
%                                                             %
%       Author: Panos Christakoglou                           %
%               Markus Oldenburg                              %
%                                                             %
%       Last changes                                          %  
%  15.10.06: Summary and abstract by Markus                   %
%  22.12.06: Added the CAF reference                          %
%  22.12.06: Changes in Introduction.tex (Fed - Yves)         %
%  22.12.06: Changes in Flow.tex (Fed - Yves - Panos)         %
%  22.12.06: Changes in Local.tex (Fed - Yves - Panos)        %
%            Changing order (example comes first)             %    
%  23.12.06: Changes in Interactive.tex (Fed - Yves - Panos)  %
%            Changing order (example comes first)             %    
%  23.12.06: Changes in Batch.tex (Fed - Yves - Panos)        %
%                                                             %    
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt,a4paper,twoside]{article}
\usepackage{indentfirst}
\usepackage{epsfig}
\usepackage[dvips]{color}
\usepackage{listings}
\lstset{                      % general command to set parameter(s)
  basicstyle=\ttfamily,       % print whole listing small
  keywordstyle=\bfseries,     % bold black keywords
  identifierstyle=,           % identifiers in italic
  commentstyle=\itshape,      % white comments in italic
  stringstyle=\ttfamily,      % typewriter type for strings
  showstringspaces=false,     % no special string spaces
  columns=fullflexible,       % Flexible columns
  xleftmargin=2em,            % Extra margin, left
  xrightmargin=2em,           % Extra margin, right
  numbers=left,               % Line numbers on the left
  numberfirstline=true,       % First line numbered
  firstnumber=1,              % Always start at 1 
  stepnumber=5,               % Every fifth line 
  numberstyle=\footnotesize\itshape, % Style of line numbers
  frame=lines}                % Lines above and below listings

\topmargin=1cm

\newcommand{\tag}{{\ttfamily Event Tag System }}

\begin{document}

\title{Distributed Analysis - Overview of the framework and practical examples}
\author{Editors: P.~Christakoglou\footnote{{\bf email : Panos.Christakoglou@cern.ch}}, M.~Oldenburg\footnote{{\bf email : Markus.Oldenburg@cern.ch}}}
\date{\today}

\maketitle

\begin{abstract}
In order to perform physics analysis in ALICE, a physicist has to use the GRID infrastructure since the data will be distributed in many sites. The machinery though complex due to the nature of the GRID, is totally transparent to the user who is shielded by a friendly user interface. In order to provide some guidelines to successfully utilize the provided functionalities, it was felt that an up-to-date manual was needed. With this document we try to explain the analysis framework as of today by covering all the recent developments.

After a short introduction, the general flow of a GRID based analysis will be outlined. The different steps of how to get your analysis ready and running on the GRID using the newly developed analysis framework will be covered in the subsequent chapters.


\end{abstract}

\newpage
\tableofcontents
%\lstlistoflistings


%________________________________________________________
\include{Introduction}
%________________________________________________________
\include{Flow}
%________________________________________________________
\include{Analysis}
%________________________________________________________
\include{Local}
%________________________________________________________
\include{Interactive}
%________________________________________________________
\include{Batch}
%________________________________________________________
\include{Summary}
%________________________________________________________
\begin{appendix}
\include{ObjectCuts}
\include{StringCuts}
\end{appendix}
%________________________________________________________


%________________________________________________________
\begin{thebibliography}{99}
%________________________________________________________
\bibitem{Note:RefPPR} {"ALICE Physics Performance Report", CERN/LHCC 2003-049.}
\bibitem{Note:RefComputingTDR} {F.\ Carminati \textit{et al.}, ALICE collaboration, CERN/LHCC 2005.}
\bibitem{Note:RefCAF} {CAF web cite.}
\bibitem{Note:RefAnalysisFramework} {http://pcaliweb02.cern.ch/Offline/Analysis/AnalysisFramework/}
\bibitem{Note:RefFileCatalogMetadataNote} {M.\ Oldenburg, ALICE internal note --- to be submitted to EDMS;\\
http://cern.ch/Oldenburg/MetaData/MetaData.doc}
\bibitem{Note:RefEventTagNote} {P.\ Christakoglou and P.\ Hristov, "The Event Tag System", ALICE-INT-2006-023.}
%________________________________________________________
\bibitem{Note:RefFileCatalogMetadataWeb} {http://pcaliweb02.cern.ch/Offline/Analysis/RunEventTagSystem/\\RunTags.html\#Run/File\%20metadata}
\bibitem{Note:RefAlienTutorial} {http://pcaliweb02.cern.ch/Offline/Analysis/Tutorial/\\
http://indico.cern.ch/conferenceDisplay.py?confId=8546}
\bibitem{Note:RefALIEN} {
P.~Saiz \textit{et al.}, Nucl.\,Instrum.\,Meth.\,A\textbf{502} (2003) 437--440;\\ 
http://alien.cern.ch/.}
\bibitem{Note:RefPROOF} {http://pcaliweb02.cern.ch/Offline/Analysis/CAF/}
%________________________________________________________
\bibitem{Note:ROOTManual} {http://root.cern.ch/root/html/TTreeFormula.html}
%________________________________________________________
\bibitem{Note:RefEventTagWeb} {http://pcaliweb02.cern.ch/Offline/Analysis/RunEventTagSystem/\\EventTagsAnalysis.html\#Analysis\%20with\%20tags}
%________________________________________________________
\bibitem{Note:RefGSHELL} {http://project-arda-dev.web.cern.ch/project-arda-dev/alice/apiservice/AA-UserGuide-0.0m.pdf}
%________________________________________________________
\end{thebibliography}
%________________________________________________________


\end{document}
