\documentclass{elsart}  
\usepackage{epsfig,amssymb,amsmath}  
\begin{document}


\section{ALICE TPC  ExB effect}

The main barrel of ALICE is located inside the L3 magnet.
The magnetic field B of the L3 magnet and the electric field E of the TPC are in the 
first approximation parallel.  The magnetic field supplied by the L3 magnet (it
 measures about length x diameter 11.6 x 11.2 m) is not perfectly homegenous.
In the region where the TPC is located these inhomogenities are on the order of 1\%.

The small angle of displacement causes a displacement of tracks that is 
proportional to the drift length and parameter $\omega\tau$.
Parameter  $\omega\tau$ is, in its turn, proportional to the magnetic field ($\omega=(e/m)|B|)$ 
and a characteristic time of TPC drift gas $\tau$. Integrating over the full 
drift length of ALICE TPC (250 cm), these displacements are of the order of 1 
cm and have to be corrected.

ExB effect in the ALICE TPC was calibrated using ALICE laser system which consists of 336 laser beams spaced at different x, y, z position. Displacement of  reconstructed laser beam position with respect to sureveyed position for different magnetic field setting was measured ($B_z$ = -0.5, -0.5,-0.2, 0, 0.2, 0.4 and 0.5 T). The effect of misalingment was reduced by subtracting the displacement values for the setting without magnetic field.

The measured L3 magnetic field can serve as additional input source for the calibration procedure.
Parameter $\omega\tau$ was fitted by measuring the laser track distortion at different magnetic field intensity . 

The precision of constructed ExB distiortion map is on the level of about 0.35 mm. Additional improvement was obtained by adding constant B field components on top of measured B field map. (Not included in the AliRoot yet)  
\end{document}
