\documentclass[11pt]{article}
\usepackage[margin=2cm,twoside,a4paper]{geometry}
\usepackage{amstext}
\usepackage{amsmath}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\newcommand{\AbbrName}[1]{{\ifmmode\text{\scshape #1}\else{\scshape #1}\fi}}
\newcommand{\SPD}{\AbbrName{spd}}
\newcommand{\ESD}{\AbbrName{esd}}
\newcommand{\AOD}{\AbbrName{aod}}
\newcommand{\INEL}{\AbbrName{inel}}
\newcommand{\INELONE}{$\AbbrName{inel}>0$}
\newcommand{\NSD}{\AbbrName{nsd}}
\newcommand{\FMD}[1][]{\AbbrName{fmd\ifx|#1|\else#1\fi}}
\newcommand{\dndetadphi}[1][]{{\ensuremath% 
    \ifx|#1|\else\left.\fi%
    \frac{d^2N_{ch}}{d\eta\,d\varphi}%
    \ifx|#1|\else\right|_{#1}\fi%
}}
\newcommand{\landau}[1]{{\ensuremath% 
    \text{landau}\left(#1\right)}}
\newcommand{\dndeta}[1][]{{\ensuremath% 
    \ifx|#1|\else\left.\fi%
    \frac{1}{N}\frac{dN_{ch}}{d\eta}%
    \ifx|#1|\else\right|_{#1}\fi%
}}
\newcommand{\Nsel}{{\ensuremath N_{v,\text{trigger,vertex}}}}
\setlength{\parskip}{1ex}
\setlength{\parindent}{0em}

\title{Analysing the FMD data for $\dndeta$}
\author{Christian Holm
  Christensen\footnote{\texttt{$\langle$cholm@nbi,dk$\rangle$}}\quad\&\quad
  Hans Hjersing Dalsgaard\footnote{\texttt{$\langle$canute@nbi,dk$\rangle$}}\\ 
  Niels Bohr Institute\\
  University of Copenhagen}
\date{\today}
\begin{document}
\maketitle 

\section*{Introduction}

This document describes the steps performed in the analysis of the
charged particle multiplicity in the forward pseudo--rapidity
regions. 

The analysis is performed as a two--step process.  
\begin{enumerate}
\item The Event--Summary--Data (\ESD{}) is processed event--by--event
  and passed through a number of algorithms, and
  $\dndetadphi$ for each event is output to an Analysis--Object--Data
  (\AOD{}) tree.
\item The \AOD{} data is read in and the sub--sample of the data under
  investigation is selected (e.g., \INEL{}, \INELONE{}, \NSD{}, or
  some centrality class) and the $\dndetadphi$ histogram read in for
  those events to build up $\dndeta$
\end{enumerate}
The details of each step above will be expanded upon in the
following. 

\section*{Generating $\dndetadphi[i]$ event--by--event}

When reading in the \ESD{}s and generating the $\dndetadphi$
event--by--event the following steps are taken (in order) for each
event $i$
\begin{description}
\item[Event inspection] The global properties of the event is
  determined, including the trigger type, vertex $z$ coordinate, and
  whether this is a low--flux event or not. 
\item[Sharing filter] The \ESD{} object is read in and corrected for
  sharing.  The result is a new \ESD{} object.
\item[Density calculator] The (possibly un--corrected) \ESD{} object
  is then inspected and an inclusive, per--ring charged particle
  density $\dndetadphi[incl,r,v,i]$  is made.  This
  calculation depends in general upon the interaction
  vertex\footnote{In the following simply labelled 'primary vertex' or
    'vertex'.} position along the $z$ axis ($v_z$).  
\item[Corrections] The 5 $\dndetadphi[incl,r,v,i]$ are
  corrected for secondary production, event selection efficiency, and
  possibly the sharing efficiency.  These corrections are highly
  dependent on the vertex $z$ coordinate.  The result is an per--ring,
  charged primary particle density $\dndetadphi[r,v,i]$
\item[Histogram collector] Finally, the 5 $\dndetadphi[r,v,i]$ are
  summed into a single $\dndetadphi[v,i]$ histogram, taking care of
  the overlaps between the detector rings.  In principle, this
  histogram is independent of the vertex, except that the
  pseudo--rapidity range, and possible holes in that range, depends on
  $v_z$ --- or rather the bin in which the $v_z$ falls. 
\end{description}

Each of these steps will be detailed in the following. 

\subsection*{Event inspection}

The first thing to do, is to inspect the event for triggers.  A number
of trigger bits, like \INEL{}, \INELONE{}, \NSD{}, and so on is then
propagated to the \AOD{} output.  

Next, the number of \emph{tracklets} reconstructed in the Silicon
Pixel Detector (\SPD{}) compared to a threshold.  If the number of
track--lets falls belows this threshold, the event is consider a
low--flux event.   

Just after the sharing filter (described below) but before any more
processing, the vertex information is queried.  If there is no vertex
information, or if the vertex $z$ coordinate is outside the
pre--defined range, then no further processing takes place. 

\subsection*{Sharing filter}

The \FMD{} \ESD{} object contains the scaled energy deposited $\Delta
E/\Delta E_{mip}$ for each of the 51,200 strips.  The \FMD{} is
organised in 3 \emph{sub--detectors} \FMD{1}, \FMD{2}, and \FMD{3}, each
consisting of 1 (\FMD{1}) or 2 (\FMD{2} and \FMD{3}) \emph{rings}.
The rings fall into two types: \emph{Inner} or \emph{outer} rings.
Each ring is in turn  azimuthal divided into \emph{sectors}, and each
sector is radially divided into \emph{strips}.  How many sectors,
strips, as well as the $\eta$ coverage is given in
\tablename~\ref{tab:fmd:overview}. 

\begin{table}[htbp]
  \begin{center}
    \caption{Physical dimensions of Si segments and strips.}
    \label{tab:fmd:overview}
    \vglue0.2cm
    \begin{tabular}{|c|cc|cr@{\space--\space}l|r@{\space--\space}l|}
      \hline
      \textbf{Sub--detector/} &
      \textbf{Azimuthal}&
      \textbf{Radial} &
      $z$ &
      \multicolumn{2}{c|}{\textbf{$r$}} &
      \multicolumn{2}{c|}{\textbf{$\eta$}} \\
      \textbf{Ring}& 
      \textbf{sectors} &
      \textbf{strips} & 
      \textbf{[cm]} &
      \multicolumn{2}{c|}{\textbf{range [cm]}} &
      \multicolumn{2}{c|}{\textbf{coverage}} \\
      \hline
      FMD1i & 20& 512& 320  &  4.2& 17.2& 3.68&  5.03\\
      FMD2i & 20& 512&  83.4&  4.2& 17.2& 2.28&  3.68\\
      FMD2o & 40& 256&  75.2& 15.4& 28.4& 1.70&  2.29\\
      FMD3i & 20& 512& -75.2&  4.2& 17.2&-2.29& -1.70\\
      FMD3o & 40& 256& -83.4& 15.4& 28.4&-3.40& -2.01\\
      \hline
    \end{tabular}
  \end{center}
\end{table}

A particle originating from the vertex can, because of it's incident
angle on the \FMD{} sensors traverse more than one strip.  That means
that the energy loss of the particle is distributed over 1 or more
strips.  The signal in each strip should therefore possibly merged
with it's neighbor strip signals to properly reconstruct the energy
loss of a single particle.  

The effect is most pronounced in low--flux events, like proton--proton
collisions or peripheral Pb--Pb collisions, while in high--flux events
the hit density is so high that most likely each and every strip will
be hit and the effect cancel out on average. 

Since the particles travel more or less in straight lines toward the
\FMD{} sensors, the sharing effect predominantly in the $r$ or
\emph{strip} direction.  Only neighboring strips in a given sector is
therefor investigated for this effect.  

Algorithm~\ref{algo:sharing} is applied to the signals in a given
sector.

\begin{algorithm}[htpb]
  \SetKwData{usedThis}{current strip used}
  \SetKwData{usedPrev}{previous strip used}
  \SetKwData{Output}{output}
  \SetKwData{Input}{input}
  \SetKwData{Nstr}{\# strips}
  \SetKwData{Signal}{current}
  \SetKwData{Eta}{$\eta$}
  \SetKwData{prevE}{previous strip signal} 
  \SetKwData{nextE}{next strip signal} 
  \SetKwData{lowFlux}{low flux flag} 
  \SetKwFunction{SignalInStrip}{SignalInStrip}
  \SetKwFunction{MultiplicityOfStrip}{MultiplicityOfStrip}
  \usedThis $\leftarrow$ false\;
  \usedPrev $\leftarrow$ false\;
  \For{$t\leftarrow1$ \KwTo \Nstr}{ 
    \Output${}_t\leftarrow 0$\;
    \Signal $\leftarrow$ \SignalInStrip($t$)\;

    \uIf{\Signal is not valid}{ 
      \Output${}_t \leftarrow$ invalid\;
    }
    \uElseIf{\Signal is 0}{ 
      \Output${}_t \leftarrow$ 0\;
    }
    \Else{
      \Eta$\leftarrow$ $\eta$ of \Input${}_t$\;
      \prevE$\leftarrow$ 0\;
      \nextE$\leftarrow$ 0\;
      \lIf{$t \ne 1$}{ 
        \prevE$\leftarrow$ \SignalInStrip($t-1$)\;
      }
      \lIf{$t \ne $\Nstr}{ 
        \nextE$\leftarrow$ \SignalInStrip($t+1$)\;
      }
      \Output${}_t\leftarrow$
      \MultiplicityOfStrip(\Signal,\Eta,\prevE,\nextE,\\
      \hfill\lowFlux,$t$,\usedPrev,\usedThis)\;
    }   
  }
  \caption{Sharing correction}
  \label{algo:sharing}
\end{algorithm}

Here the function \FuncSty{SignalInStrip}($t$) returns the properly
path--length corrected signal in strip $t$.  The function
\FuncSty{MultiplicityInStrip} is where the real processing takes
place (see page \pageref{func:MultiplicityInStrip}). 

\begin{function}[htbp]
  \caption{MultiplicityInStrip(\DataSty{current},$\eta$,\DataSty{previous},\DataSty{next},\DataSty{low
      flux flag},\DataSty{previous signal used},\DataSty{this signal
      used})} 
  \label{func:MultiplicityInStrip}
  \SetKwData{Current}{current} 
  \SetKwData{Next}{next} 
  \SetKwData{Previous}{previous} 
  \SetKwData{lowFlux}{low flux flag}
  \SetKwData{usedPrev}{previous signal used}
  \SetKwData{usedThis}{this signal used}
  \SetKwData{lowCut}{low cut}
  \SetKwData{total}{Total}
  \SetKwData{highCut}{high cut}
  \SetKwData{Eta}{$\eta$}  
  \SetKwFunction{GetHighCut}{GetHighCut}
  \If{\Current is very large or \Current $<$ \lowCut} {
    \usedThis $\leftarrow$ false\;
    \usedPrev $\leftarrow$ false\;
    \Return{0}
  }
  \If{\usedThis}{ 
    \usedThis $\leftarrow$ false\;
    \usedPrev $\leftarrow$ true\;
    \Return{0}
  }
  \highCut $\leftarrow$ \GetHighCut($t$,\Eta)\;
  \If{\Current $<$ \Next and \Next $>$ \highCut and \lowFlux set}{ 
    \usedThis $\leftarrow$ false\;
    \usedPrev $\leftarrow$ false\;
    \Return{0}
  }
  \total $\leftarrow$ \Current\;
  \lIf{\lowCut $<$ \Previous $<$ \highCut and not \usedPrev}{ 
    \total $\leftarrow$ \total + \Previous\;
  }
  \If{\lowCut $<$ \Next $<$ \highCut}{ 
    \total $\leftarrow$ \total + \Next\;  
    \usedThis $\leftarrow$ true\;
  }
  \eIf{\total $>$ 0}{ 
    \usedPrev $\leftarrow$ true\;
    \Return{\total}
  }{
    \usedPrev $\leftarrow$ false\;
    \usedThis $\leftarrow$ false\;
    \Return{0}
  }
\end{function}
Here, the function \FuncSty{GetHighCut} evaluates a Landau
distribution fitted to the energy spectrum in the $\eta$ bin
specified.  It returns 
$$
\Delta_{mp} - 2 w
$$
where $\Delta_{mp}$ is the most probable energy loss, and $w$ is the
width of the Landau distribution.  

The \KwSty{if} in line 5, says that if the previous strip was merged
with current one, and the signal of the current strip was added to
that, then we the current signal is set to 0, and we mark it as used
for the next iteration (\DataSty{previous signal
  used}$\leftarrow$true). 

The \KwSty{if} in line 10 checks if the current signal is smaller than
the next signal, the next signal is larger than the upper cut defined
above, and if we have a low--flux event.  If that condition is met,
then the current signal is the smaller of two possible candidates for
merging, and it should be merged into the next signal.  Note, that
this \emph{only} applies in low--flux events.  

On line 15, we test if the previous signal lies between our low and
high cuts, and if it has not been marked as being used.  If so, we add
it to our current signal.  

The next \KwSty{if} on line 16 checks if the next signal is within our
cut bounds.  If so, we add that signal to the current signal and mark
it as used for the next iteration (\DataSty{this signal
  used}$\leftarrow$true).  It will then be zero'ed on the next
iteration by the condition on line 6.

Finally, if our signal is still larger than 0, we return the signal
and mark this signal as used (\DataSty{previous signal
  used}$\leftarrow$true) so that it will not be used in the next
iteration. Otherwise, we mark the current signal and the next signal
as unused and return a 0. 


\subsection*{Density calculator}

The density calculator loops over all the strip signals and calculates
the inclusive (primary + secondary) charged particle density in
pre-defined $(\eta,\varphi)$ bins.  

If the event is classified as a low--flux event, then the number of
charged particles in a given by a simple threshold: 
\begin{align}
  N_{ch,t} &= \left\{
    \begin{array}{cl}
      0 & \Delta_t < \text{low cut}\\ 
      1 & \Delta_t \ge \text{low cut}\\ 
    \end{array}\right.\quad,
\end{align}
where $t$ is the strip identifier, $\Delta_t$ is the scaled energy
deposition in that strip, and 'low cut' is a predefined cut.  For high
flux events, the number charged particles in a strip is calculated
using multiple Landau distributions fitted to the energy loss spectrum
at a given $\eta$ value.  
\begin{align}
  \Delta_{2,mp} &= 2 \Delta_{mp}+ 2 w \log(2)\nonumber\\
  \Delta_{3,mp} &= 3 \Delta_{mp}+ 3 w \log(2)\nonumber\\
  N_{ch,t} &= \frac{\landau{\Delta_t,\Delta_{mp},w}+
    2\,\alpha\,\landau{\Delta_t,\Delta_{2,mp},2w} + 
    3\,\beta\,\landau{\Delta_t,\Delta_{3,mp},3w}}{% 
    \landau{\Delta_t,\Delta_{mp},w}+
    \alpha\,\landau{\Delta_t,\Delta_{2,mp},2w} + 
    \beta\,\landau{\Delta_t,\Delta_{3,mp},3w}}\quad,
\end{align}
where $\landau{x,\psi,W}$ is the evaluation of the Landau distribution
with most probable value $\psi$ and width $W$ at $x$, $w$ is the width
of the first MIP peak, $\Delta_{mp}$ the most probable value of
the first MIP peak, and $\alpha$ and
$\beta$ are the relative strength of the second and third MIP peak in
the fitted energy loss spectrum. 

But before the signal $N_{ch,t}$ can be added to the $(\eta,\varphi)$
bin in one of the 5 per--ring histograms, it needs to be corrected for
the $\varphi$ acceptance of the strip, as well as a correction for
double hits in low--flux events.   

The acceptance correction is only applicable where the strip length
does not cover the full sector.  This is the case for the outer strips
in both the inner and outer type rings.  The acceptance correction is
then simply 
\begin{align}
  \label{eq:acc_corr}
  a_t &= \frac{l_t}{\Delta\varphi}\quad
\end{align}
where $l_t$ is the strip length in radians at constant $r$, and
$\Delta\varphi$ is $2\pi$ divided by the number of sectors in the
ring (20 for inner type rings, and 40 for outer type rings). 

Even in low--flux events it is possible that more than one particle
hits a strip.  However, for low--flux events, it is not possible to
reconstruct the 3\textsuperscript{rd} nor even the
2\textsuperscript{nd} MIP peak in the energy loss spectrum.
Therefore, the strip signal needs to be corrected to the average
number of particle impinging on a strip at a given $\eta$.  
\begin{align}
  d_t &= \left\{\begin{array}{cl} \langle n_t\rangle & \text{low
        flux}\\
      1 & \text{high flux}
    \end{array}\right.
\end{align}

The final $(\eta,\varphi)$ content of the 5 output vertex dependent,
per--ring histograms of the inclusive charged particle density is then
given by
\begin{align}
  \label{eq:density}
  \dndetadphi[incl,r,v,i(\eta,\varphi)] &= \sum_t^{t\in(\eta,\varphi)}
  N_{ch,t}\,a_t\,d_t
\end{align}
where $t$ runs over the strips in the $(\eta,\varphi)$ bin. 

\subsection*{Corrections}

The corrections code receives the five vertex dependent,
per--ring histograms of the inclusive charged particle density
$\dndetadphi[incl,r,v,i]$ from the density calculator and applies
three corrections 
\begin{description}
\item[Secondary correction:] This is a 2 dimensional histogram
  generated from simulations of the ratio of primary particles to the
  total number of particles that fall within an $(\eta,\varphi)$ bin
  for a given vertex bin

  \begin{align}
    s_v(\eta,\varphi) &=
    \frac{\sum_i^{N_{v}}N_{ch,\text{primary},i}(\eta,\varphi)}{
      \sum_i^{N_{v}}N_{ch,\text{\FMD{}},i}(\eta,\varphi)}\quad,
  \end{align}
  where $N_{v}$ is the number of events with a vertex bin $v$, and
  $N_{ch,\FMD{},i}$ is the total number of charged particles that hit
  the \FMD{} in event $i$ in the specified $(\eta,\varphi)$ bin and
  $N_{ch,\text{primary},i}$ is number of primary charged particles in
  event $i$ within the specified $(\eta,\varphi)$ bin.

  $N_{ch,primary}(\eta,\varphi)$ is given by summing over the charged
  particles labelled as primaries \emph{at the time of the collision}
  as defined in the simulation code.  That is, it is the number of
  primaries within the bin at the collision point --- not at the
  \FMD{}.  
\item[Event selection efficiency:] 
%% AliFMDAnalysisTaskGenerateCorrection
%% 
%% Analysed_FMD<r>_vtx<v> 2D histogram (eta,phi)
%% EventsSelected[<v>]    bin content 
%% Inel_FMD<r>_vtx<v>     2D histogram (eta,phi) 
%% EventsAll[<v>]         bin content 
%%
%% correction is 
%%
%%      Analysed_FMD<d><r>_vtx<v> / EventsSelected[<v>]
%%    ---------------------------------------------------
%%      Inel_FMD<r>_vtx<v> / EventsAll[<v>] 
%% 
%% Inel_FMD<r>_vtx<v>        is d^2N_incl,ch/d\eta d\varphi for all
%%                           events  
%% Analysed_FMD<d><r>_vtx<v> is d^2N_{incl,ch}/d\eta d\varphi for
%%                           events with trigger _and_ vertex
%% EventsSelected            is dN/dv_z for events with trigger _and_
%%                           vertex
%% EventsAll                 is dN/dv_z for all events 
  This correction is made from simulations.  It is a 2--dimensional
  histogram of the ratio of average number of charged particles in
  events that are triggered as interactions \emph{and} have vertex, to
  the average number of charged particles in all events in
  $(\eta,\varphi)$ bins
  \begin{align}
    \label{eq:event_sel_eff}
    e_{v}(\eta,\varphi) &= 
    \frac{\frac{1}{\Nsel{}}
      \sum_{i}^{\Nsel{}} N_{ch,i}(\eta,\varphi)}{
      \frac{1}{N_{v}}
      \sum_{i}^{N_{v}} N_{ch,i}(\eta,\varphi)}\quad,
  \end{align}
  where $N_{v}$ is the total number of events in vertex bin $v$, and
  $\Nsel{}$ is the number of events with a collision trigger and
  vertex within the predefined range, for vertex bin $v$.  $N_{ch,i}$
  is the number of charged 'MC truth' particles within the specified
  $(\eta,\varphi)$ range for event $i$.

\item[Sharing correction efficiency:] 
%% hits_NoCuts_FMD<d><r>_vtxbin<v>_proj 
%% hMCHits_nocuts_FMD<d><r>_vtxbin<v>
%% 
%% - hits_NoCuts_FMD<d><r>_vtxbin<v>_proj 
%%   A projection of hits_FMD<d><r>_vtxbin<v> onto the eta axis and
%%   possibly scaled to EventSel1D. 
%% - EventSel1D - the 1D event selection efficiency (per vertex) 
%%   This is the ratio of hEventAll divided hEventsSelected, with
%%   errors calculated as 
%% 
%%      da = sqrt(e_sel^2 + e_all^2)
%%      a  = c_all - c_sel 
%%      e  = sqrt(da^2 + a^2 / (c_ell*c_sel)^2)
%%         = sqrt(e_sel^2 + e_all^2 + (c_all-c_sel)^2 /
%%         (c_ell*c_sel)^2)
%%
%%   where hEventsAll[<v>] is the total number of events in the v bin,
%%   and hEventsSelected[<v>] is the notoal number of events with
%%   trigger and vertex in the v bin. Both for MC
%% - hits_FMD<d><r>_vtxbin<v>
%%   from AliFMDAnalysisTaskBackgroundCorrection - the uncorrected
%%   multiplicity in each ring - i.e.,
%%   \dndetadphi[incl,r,v,i(\eta,\varphi)] 
%% - hits_MC_FMD<d><r>_vtxbin<v> - from AliFMDAnalysisTaskSharing - a 1D
%%   histogram in eta - sum of eta of track-refs. 
%% - hEvents[<v>] - bin content of dN/dv_{z} 
%% - nMCEventsNoCuts -  hPrimEvents[<v>] - bin content of dN/dv_{z}
%%   for MC events - from AliFMDAnalysisTaskSharing - number of not
%%   empty mc events per vertex bin 
%% 
  This is a one dimension histogram in $\eta$, generated from running
  the full analysis on simulated data.   
  \begin{align}
    \label{eq:sharing_corr}
    m_v(\eta) &= \frac{
      \frac{1}{e_{v}}\frac1{\Nsel{}}\left.\frac{dN_{ch}}{d\eta}\right|_{incl,r,v(\eta)}}{
      \frac1{\Nsel}\sum_i^{\Nsel{}}N_{ch,\text{MC},i(\eta)}}
    \nonumber\\
    &=
    \frac{\frac1{N_v}\left.\frac{dN_{ch}}{d\eta}\right|_{incl,r,v(\eta)}}{
      \frac1{\Nsel}\sum_i^{\Nsel{}}N_{ch,\text{MC},i(\eta)}}\nonumber\\
    &= 
    \frac{\Nsel{}}{N_v}
    \frac{\left.\frac{dN_{ch}}{d\eta}\right|_{incl,r,v(\eta)}}{
      \sum_i^{\Nsel{}}N_{ch,\text{MC},i(\eta)}}
  \end{align}
  since
  \begin{align}
    e_{v} &= \frac{N_{v}}{\Nsel{}}\quad,
%% \pm 
%%    \sqrt{\sqrt{\Nsel{}}^2 + \sqrt{N_v^2} +
%%      \frac{(N_v-\Nsel{})^2}{(\Nsel{}N_v)^2}}\nonumber \\
%%    &= \frac{N_{v}}{\Nsel{}} \pm 
%%    \sqrt{\Nsel{} + N_v + \frac1{\Nsel{}^2} + \frac1{N_v^2} -
%%      \frac1{\Nsel{}N_v}}\quad, 
  \end{align}
  and $N_v$ is the total number of simulated events with vertex in bin
  $v$, $\Nsel{}$ is the number of simulated events with a collision
  trigger and a reconstructed vertex, $N_{ch,\text{MC},i}(\eta)$ is
  the total number of 'MC truth' charged particles in the given eta
  bin in event $i$, and 
  \begin{align}
    \left.\frac{dN_{ch}}{d\eta}\right|_{incl,r,v(\eta)} &= 
    \int_0^{2\pi}d\varphi\dndetadphi[incl,r,v(\eta,\varphi)]
    \nonumber\\
    &=  \int_0^{2\pi}d\varphi\sum_i^{\Nsel{}}\dndetadphi[incl,i,r,v(\eta,\varphi)]\quad,
  \end{align}
  where the integrand is \eqref{eq:density} for simulated data i.e.,
  this is the inclusive total charge particle density in a given $v$
  bin for ring $r$ in event $i$. 
\end{description}

The 5 output vertex dependent, per--ring histograms of the primary
charged particle density is then given by
\begin{align}
  \dndetadphi[r,v,i(\eta,\varphi)] &=
  s_v(\eta,\varphi)\,e_v(\eta,\varphi)\,m_v(\eta)\dndetadphi[incl,r,v,i(\eta,\varphi)]
\end{align}

\subsection*{Histogram collector}

The histogram collector collects the information from the 5 vertex
dependent, per--ring histograms of the primary charged particle
density $\dndetadphi[r,v,i]$ into a single vertex dependent histogram
of the charged particle density $\dndetadphi[v,i]$.  

To do this, it first calculates, for each vertex bin, the $\eta$ bin
range to use for each ring.  It investigates the secondary correction
maps $s_v(\eta,\varphi)$ to find the edges of the map.  The edges are
given by the $\eta$ range where $s_v(\eta,\varphi)$ is larger than
some threshold\footnote{Typically $t_s\approx 0.1$.}  $t_s$. The code
applies safety margin of a $N_{cut}$ bins\footnote{Typically
  $N_{cut}=1$.}, to ensure that the data selected does not have too
large corrections associated with it.

It then loops over the bins in the defined $\eta$ range and sums the
contributions from each of the 5 histograms.  In the $\eta$ ranges
where two rings overlap, the collector calculates the average and adds
the errors in quadrature.

The output vertex dependent histogram of the primary
charged particle density is then given by
\begin{align}
  \dndetadphi[v,i(\eta,\varphi)] &=
  \frac{1}{N_{r\in(\eta,\varphi)}}\sum_{r}^{r\in(\eta,\varphi)}  
  \dndetadphi[r,v,i(\eta,\varphi)]\\
  \delta\left[\dndetadphi[v,i(\eta,\varphi)]\right] &=
  \frac{1}{N_{r\in(\eta,\varphi)}}\sqrt{\sum_{r}^{r\in(\eta,\varphi)}   
    \delta\left[\dndetadphi[r,v,i(\eta,\varphi)]\right]^2}\\
  \quad,
\end{align}
where $N_{r\in(\eta,\varphi)}$ is the number of overlapping histograms
in the given $(\eta,\varphi)$ bin. 


\section*{Building the final $\dndeta$}



\end{document}
